---
title: "Part 2: Analyzing DADA2 Output With Phyloseq"
author: "David Ramotowski"
date: "12/22/2022"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Part 4: Further Microbiome Analysis Using Phyloseq
  
  Upload sample metadata file (optional, good if comparing multiple factors)
    -Make sure sample names in 1st column match names of the otu table
```{r}
  samdf <- data.frame(read.csv('sample_meta.csv', header = TRUE, 
                               row.names = 1, sep = ","))
  head(samdf)
```

  Create Phyloseq object 
  (remove sample_data command if you are not using a metadata file)  
```{r}
  theme_set(theme_bw())
  ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE), 
                 tax_table(taxa), sample_data(samdf)) 
  #creates phyloseq object from DADA2 outputs
  dna <- Biostrings::DNAStringSet(taxa_names(ps))
  names(dna) <- taxa_names(ps)
  ps <- merge_phyloseq(ps, dna)
  taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))
  ps
```

    Create ASV and Taxonomy Tables
```{r}
write.table(as.data.table(otu_table(ps), keep.rownames = T), file = 
              "expr.otu_table.tsv", 
        sep = "\t", quote = F, row.names = F, col.names = T)
write.table(as.data.table(tax_table(ps), keep.rownames = T), file = 
              "expr.tax_table.tsv", 
        sep = "\t", quote = F, row.names = F, col.names = T)
```

  Create raw abundance tables 
    Remove Sample_ID in group_by() it doesn't exist, then replace 
    pivot_wider(names_from = Sample_ID, values_from = Abundance) with 
    spread(Sample_ID, Abundance), you can also change Sample_ID to other 
    variables from metadata.
```{r}
ifelse(!dir.exists("raw_abundance_tables"), dir.create("raw_abundance_tables"), 
       FALSE)
write.table(ps %>% psmelt() %>% 
        arrange(OTU) %>% dplyr::rename(ASV = OTU) %>% 
        dplyr::select(ASV, Kingdom, Phylum, Class, Order, Family, Genus, 
               Sample_ID, Abundance) %>%
        pivot_wider(names_from = Sample_ID, values_from = Abundance), 
    file = "raw_abundance_tables/expr.abundance.all.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
write.table(ps %>% tax_glom(taxrank = "Phylum", bad_empty = c("", " ", "\t")) %>% psmelt() %>% dplyr::select(Phylum, Sample_ID, Abundance) %>% dplyr::group_by(Phylum, Sample_ID) 
        %>% dplyr::summarise(Abundance = sum(Abundance)) %>% 
        pivot_wider(names_from = Sample_ID, values_from = Abundance), 
    file = "raw_abundance_tables/expr.abundance.abphy.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
write.table(ps %>% tax_glom(taxrank = "Class", bad_empty = c("", " ", "\t")) %>% psmelt() %>%    dplyr::select(Class, Sample_ID, Abundance) %>% dplyr::group_by(Class, Sample_ID) 
        %>% dplyr::summarise(Abundance = sum(Abundance)) %>% 
        pivot_wider(names_from = Sample_ID, values_from = Abundance), 
    file = "raw_abundance_tables/expr.abundance.abcls.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
write.table(ps %>% tax_glom(taxrank = "Family" , bad_empty = c("", " ", "\t")) %>% psmelt() %>%   dplyr::select(Family, Sample_ID, Abundance) %>% dplyr::group_by(Family, Sample_ID) 
        %>% dplyr::summarise(Abundance = sum(Abundance)) %>% 
        pivot_wider(names_from = Sample_ID, values_from = Abundance), 
    file = "raw_abundance_tables/expr.abundance.abfam.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
write.table(ps %>% tax_glom(taxrank = "Genus", bad_empty = c("", " ", "\t")) %>% psmelt() %>%
        dplyr::select(Genus, Sample_ID, Abundance) %>% dplyr::group_by(Genus, Sample_ID) 
        %>% dplyr::summarise(Abundance = sum(Abundance)) %>% 
        pivot_wider(names_from = Sample_ID, values_from = Abundance), 
    file = "raw_abundance_tables/expr.abundance.abgen.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
```

  Create relative abundance tables 
    Remove Sample_ID in group_by() it doesn't exist, then replace 
    pivot_wider(names_from = Sample_ID, values_from = Abundance) with 
    spread(Sample_ID, Abundance), you can also change Sample_ID to other 
    variables from metadata.
```{r}
ifelse(!dir.exists("relative_abundance_tables"), 
       dir.create("relative_abundance_tables"), FALSE)

write.table(ps %>% transform_sample_counts(function(x) {x/sum(x)} ) %>% 
              psmelt() %>% 
        arrange(OTU) %>% dplyr::rename(ASV = OTU) %>% 
        dplyr::select(ASV, Kingdom, Phylum, Class, Order, Family, Genus, 
               Sample_ID, Abundance) %>%
        pivot_wider(names_from = Sample_ID, values_from = Abundance), 
    file = "relative_abundance_tables/expr.relative_abundance.all.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
write.table(ps %>% tax_glom(taxrank = "Phylum", bad_empty = c("", " ", "\t")) %>% 
        transform_sample_counts(function(x) {x/sum(x)} ) %>% psmelt() %>%   
        dplyr::select(Phylum, Sample_ID, Abundance) %>% dplyr::group_by(Phylum, Sample_ID) 
        %>% dplyr::summarise(Abundance = sum(Abundance)) %>% 
        pivot_wider(names_from = Sample_ID, values_from = Abundance), 
    file = "relative_abundance_tables/expr.relative_abundance.abphy.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
write.table(ps %>% tax_glom(taxrank = "Class", bad_empty = c("", " ", "\t")) %>% 
        transform_sample_counts(function(x) {x/sum(x)} ) %>% psmelt() %>%   
        dplyr::select(Class, Sample_ID, Abundance) %>% dplyr::group_by(Class, Sample_ID) 
        %>% dplyr::summarise(Abundance = sum(Abundance)) %>% 
        pivot_wider(names_from = Sample_ID, values_from = Abundance), 
    file = "relative_abundance_tables/expr.relative_abundance.abcls.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
write.table(ps %>% tax_glom(taxrank = "Family", bad_empty = c("", " ", "\t")) %>% 
        transform_sample_counts(function(x) {x/sum(x)} ) %>% psmelt() %>%   
        dplyr::select(Family, Sample_ID, Abundance) %>% dplyr::group_by(Family, Sample_ID) 
        %>% dplyr::summarise(Abundance = sum(Abundance)) %>% 
        pivot_wider(names_from = Sample_ID, values_from = Abundance), 
    file = "relative_abundance_tables/expr.relative_abundance.abfam.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
write.table(ps %>% tax_glom(taxrank = "Genus", bad_empty = c("", " ", "\t")) %>% 
        transform_sample_counts(function(x) {x/sum(x)} ) %>% psmelt() %>%   
        dplyr::select(Genus, Sample_ID, Abundance) %>% dplyr::group_by(Genus, Sample_ID) 
        %>% dplyr::summarise(Abundance = sum(Abundance)) %>% 
        pivot_wider(names_from = Sample_ID, values_from = Abundance), 
    file = "relative_abundance_tables/expr.relative_abundance.abgen.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
```

  Plot abundance step 1 
    Taxonomic Filtering (only used for bar plots):
```{r}
#Show available ranks in the dataset
rank_names(ps)
```

```{r}
#Create table with number of phylum features
table(tax_table(ps)[, "Phylum"], exclude = NULL)
```

  In this case, I filtered out phyla with only one feature as well as any NA
  phyla (likely artifacts), this can/will change based on your samples
```{r}
ps0 <- subset_taxa(ps, !is.na(Phylum) & !Phylum %in% 
                     c("", "Edwardsbacteria", "Hydrothermae"))
```
  
  Prevalence filtering:
```{r}
# Compute prevalence of each feature, store as data frame
prevdf <- apply(X = otu_table(ps),
               MARGIN = ifelse(taxa_are_rows(ps), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})

# Add taxonomy and total read counts to this data frame
prevdf <- data.frame(Prevalence = prevdf,
                    TotalAbundance = taxa_sums(ps),
                    tax_table(ps))

#Compute average (column 1) total prevalence (column 2) of each feature
plyr::ddply(prevdf, "Phylum", function(df1){cbind(mean(df1$Prevalence),
                                                  sum(df1$Prevalence))})
```

  Only run this chunk if there are phyla in the previous table with really low
  prevalence
```{r}
# Define phyla to filter if needed, won't filter at this step in my case
filterPhyla <- c("")

# Filter entries with unidentified Phylum.  Only run this code if you filtered 
ps0 <- subset_taxa(ps0, !Phylum %in% filterPhyla)
ps
```

  Subset and plot remaining ASV's with line showing 5% abundance threshold
  (if it's hard to see the plot, copy and run the code in the console below,
  click on the plots tab on the bottom right of RStudio and zoom)
```{r}
# Subset to the remaining phyla
prevdf1 <- subset(prevdf, Phylum %in% get_taxa_unique(ps0, "Phylum"))

# Plot
ggplot(prevdf1, aes(TotalAbundance, Prevalence / nsamples(ps0),color=Phylum)) +
  # Include a guess for parameter
  geom_point(size = 2, alpha = 0.7) +
  geom_hline(yintercept = 0.05, alpha = 0.5, linetype = 2) +
  scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") +
  facet_wrap(~Phylum) + theme(legend.position="none")
```

  Filter taxa (can change the abundance and prevalence threshold for more or 
  less taxa)
```{r}
# Define prevalence threshold as 50% of total samples
prevalenceThreshold = 0.5 * nsamples(ps0)
abundanceThreshold = 10

# Execute prevalence filter, using `prune_taxa()` function
keepTaxa <- rownames(prevdf1)[(prevdf1$Prevalence >= prevalenceThreshold 
                               & prevdf1$TotalAbundance >= abundanceThreshold)]
ps1 <- prune_taxa(keepTaxa, ps0)
```
  
  Save results to file
```{r}
saveRDS(ps1, "ps1.rds")
```
  
  Plot abundance step 2:
    calculate relative abundance
```{r}
  # Transform to proportions (relative abundances)
ps.rp = transform_sample_counts(ps1, function(OTU) OTU/sum(OTU)) 

# Top N taxa 
N = 50 #(can change to any number of taxa)
topN = names(sort(taxa_sums(ps1), decreasing=TRUE))[1:N]
ps.topN = transform_sample_counts(ps1, function(OTU) OTU/sum(OTU))
ps.topN = prune_taxa(topN, ps.topN) 

ps.topN
```
 
 Group by phylum, place low abundance samples (<1% abundance) into "other"
```{r}
glom <- tax_glom(ps.rp, taxrank = "Phylum") %>% psmelt()
glom$Phylum <- as.character(glom$Phylum)
medians <- ddply(glom, ~Phylum, function(x) c(median=median(x$Abundance)))
Other <- medians[medians$median <= 0.01,]$Phylum
glom[glom$Phylum %in% Other,]$Phylum <- 'Other'
glom <- glom %>% 
    rowwise() %>% 
    filter(Abundance != 0) %>% 
    group_by(Sample, Location, Sample_ID, Phylum) %>% 
    dplyr::summarise(Abundance = sum(Abundance))
```
 
 Plot phylum barplot
```{r}
ptaxa_trimmed = ggplot(glom, aes(x=Sample_ID, y=Abundance, fill = Phylum,)) + 
  geom_col(size = 0.05, color = "black") + facet_wrap(~Location, scales = "free_x", nrow = 1) + 
  guides(fill = guide_legend(ncol = 1)) + scale_y_continuous(breaks = seq(0, 1, 0.1)) + 
    theme(legend.title = element_text(size = 8), legend.text = element_text(size = 5), 
          axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1), 
          axis.text.y = element_text(size = 12)) + xlab("Sample") + ylab("Relative abundance") +
          ggtitle("Phylum Relative Abundance") + scale_fill_brewer(palette = "Spectral")
ptaxa_trimmed
```
I will work on genus plots later...

  Plot phylum relative abundance (older code)
```{r}
# All taxa
ifelse(!dir.exists("bar_plots"), dir.create("bar_plots"), 
       FALSE)

ptaxa1 = plot_bar(ps.rp %>% tax_glom(taxrank = "Phylum"), x = "Sample_ID", fill = "Phylum", title = paste(ntaxa(ps.rp), "Taxa (Phylum)")) + 
    geom_bar(stat = "identity", size = 0.05, color = "black") + 
    facet_wrap(~Location, scales = "free_x", nrow = 1) + guides(fill = guide_legend(ncol = 2)) + 
    scale_y_continuous(breaks = seq(0, 1, 0.1)) + 
    theme(legend.title = element_text(size = 8), legend.text = element_text(size = 5), 
          axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1), 
          axis.text.y = element_text(size = 12)) + xlab("Sample") + ylab("Relative abundance")

# Top 200 taxa
ptaxa2 = plot_bar(ps.topN %>% tax_glom(taxrank = "Phylum"), x = "Sample_ID", fill = "Phylum", title = paste(ntaxa(ps.topN), "Taxa (Phylum)")) + 
    geom_bar(stat = "identity", size = 0.05, color = "black") + 
    facet_wrap(~Location, scales = "free_x", nrow = 1) + guides(fill = guide_legend(ncol = 2)) + 
    scale_y_continuous(breaks = seq(0, 1, 0.1)) + 
    theme(legend.title = element_text(size = 8), legend.text = element_text(size = 5), 
          axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1), 
          axis.text.y = element_text(size = 12)) + xlab("Sample") + ylab("Relative abundance")

png("bar_plots/plot_bar_phylum.png", width = 10, height = 10, units = "in", res = 300)
gridExtra::grid.arrange(ptaxa1, ptaxa2, nrow = 2)
invisible(dev.off())
```

  Plot class relative abundance
```{r}
ifelse(!dir.exists("bar_plots"), dir.create("bar_plots"), 
       FALSE)
#All taxa
ptaxa3 = plot_bar(ps.rp %>% tax_glom(taxrank = "Class"), x = "Sample_ID", fill = "Class", 
          title = paste(ntaxa(ps.rp), "Taxa (Class)")) +
        geom_bar(stat = "identity", size = 0.05, color = "black") +
        facet_wrap(~Location, scales = "free_x", nrow = 1) + guides(fill = guide_legend(ncol = 5)) +
        scale_y_continuous(breaks = seq(0, 1, 0.1)) +
        theme(legend.title = element_text(size = 8), legend.text = element_text(size = 5),
              axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1),
              axis.text.y = element_text(size = 12)) + xlab("Sample") + ylab("Relative abundance")
#Top 200 taxa
ptaxa4 = plot_bar(ps.topN %>% tax_glom(taxrank = "Class"), x = "Sample_ID", fill = "Class", 
          title = paste("Top",N, "Taxa (Class)")) +
        geom_bar(stat = "identity", size = 0.1, color = "black") +
        facet_wrap(~Location, scales = "free_x", nrow = 1) + guides(fill = guide_legend(ncol = 5)) +
        scale_y_continuous(breaks = seq(0, 1, 0.1)) +
        theme(legend.title = element_text(size = 8), legend.text = element_text(size = 5),
              axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1),
              axis.text.y = element_text(size = 12)) + xlab("Sample") + ylab("Relative Abundance")

png("bar_plots/plot_bar_class.png", width = 10, height = 8, units = "in", res = 300)
gridExtra::grid.arrange(ptaxa3, ptaxa4, nrow = 2)
invisible(dev.off())
```
  
  
  Plot family relative abundance
```{r}
ifelse(!dir.exists("bar_plots"), dir.create("bar_plots"), 
       FALSE)
#Top 200 taxa
ptaxa5 = plot_bar(ps.topN %>% tax_glom(taxrank = "Family"), x = "Sample_ID", fill = "Family", 
          title = paste("Top",N, "Taxa (Family)")) +
        geom_bar(stat = "identity", size = 0.1, color = "black") +
        facet_wrap(~Location, scales = "free_x", nrow = 1) + guides(fill = guide_legend(ncol = 5)) +
        scale_y_continuous(breaks = seq(0, 1, 0.1)) +
        theme(legend.title = element_text(size = 8), legend.text = element_text(size = 5),
              axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1),
              axis.text.y = element_text(size = 12)) + xlab("Sample") + ylab("Relative Abundance")

png("bar_plots/plot_bar_family.png", width = 10, height = 8, units = "in", res = 300)
gridExtra::grid.arrange(ptaxa5)
invisible(dev.off())
```
 
  Plot genus relative abundance
```{r}
ifelse(!dir.exists("bar_plots"), dir.create("bar_plots"), 
       FALSE)
#Top 200 taxa
ptaxa6 = plot_bar(ps.topN %>% tax_glom(taxrank = "Genus"), x = "Sample_ID", fill = "Genus", 
          title = paste("Top",N, "Taxa (Genus)")) +
        geom_bar(stat = "identity", size = 0.1, color = "black") +
        facet_wrap(~Location, scales = "free_x", nrow = 1) + guides(fill = guide_legend(ncol = 5)) +
        scale_y_continuous(breaks = seq(0, 1, 0.1)) +
        theme(legend.title = element_text(size = 8), legend.text = element_text(size = 5),
              axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1),
              axis.text.y = element_text(size = 12)) + xlab("Sample") + ylab("Relative Abundance")

png("bar_plots/plot_bar_genus.png", width = 10, height = 8, units = "in", res = 300)
gridExtra::grid.arrange(ptaxa6)
invisible(dev.off())
```

  Alpha diversity (Shannon and Simpson)
    Will use the un-trimmed samples (ps) for these plots to ensure 
    accurate species richness estimates, you can modify the x and color values 
    based on your experiment
```{r}
# Select alpha-diversity measures
divIdx = c("Observed", "Chao1", "Shannon", "Simpson")

ifelse(!dir.exists("diversity_plots"), dir.create("diversity_plots"), 
       FALSE)

png("diversity_plots/plot_richness.png", width = 6, height = 5, units = "in", res = 300)
plot_richness(ps, x = "Sample_ID", measures = divIdx, color = "Location", nrow = 1) + 
    geom_point(size = 0.8) + theme_bw() + 
    theme(legend.position = "none", 
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
    labs(x = "Sample", y = "Alpha Diversity Measure")
invisible(dev.off())
```

  Alpha diversity with statistical analysis
```{r}
ifelse(!dir.exists("diversity_plots"), dir.create("diversity_plots"), 
       FALSE)

png("diversity_plots/plot_richness.png", width = 5, height = 5, units = "in", res = 300)

a_my_comparisons <- list( c("AVL", "NBH"))
symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns"))

plot_richness(ps, x="Location", measures=c("Shannon", "Simpson", "Chao1"), color = "Location")+
  geom_boxplot(alpha=0.6, width = 0.5)+ 
  theme(legend.position= "none") + theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1,size=12))+
  stat_compare_means(method = "wilcox.test", comparisons = a_my_comparisons, label = "p.signif", symnum.args = symnum.args) +
  theme_bw() + labs(x = "", y = "Alpha Diversity")
```
  
  Currently working on Beta Diversity...


  Create ASV and Taxonomy Tables with filtered taxa
```{r}
write.table(as.data.table(otu_table(ps1), keep.rownames = T), file = 
              "expr.otu_table_filtered.tsv", 
        sep = "\t", quote = F, row.names = F, col.names = T)
write.table(as.data.table(tax_table(ps1), keep.rownames = T), file = 
              "expr.tax_table_filtered.tsv", 
        sep = "\t", quote = F, row.names = F, col.names = T)
```

  Create raw abundance tables from filtered taxa 
  (Remove Sample_ID if it doesn't exist, 
  can also change to other variables from metadata)
```{r}
ifelse(!dir.exists("raw_abundance_tables_filtered"), 
       dir.create("raw_abundance_tables_filtered"), FALSE)
write.table(ps1 %>% psmelt() %>% arrange(OTU) %>% rename(ASV = OTU) %>% 
        select(ASV, Kingdom, Phylum, Class, Order, Family, Genus, Sample_ID, 
               Abundance) %>% 
        spread(Sample_ID, Abundance),
        file = "raw_abundance_tables/expr.abundance.all.tsv", sep = "\t", 
        quote = F, row.names = F, col.names = T)
write.table(ps1 %>% tax_glom(taxrank = "Phylum") %>% psmelt() %>% 
        select(Phylum, Sample_ID, Abundance) %>% spread(Sample_ID, Abundance),
        file = "raw_abundance_tables/expr.abundance.abphy.tsv", sep = "\t", 
        quote = F, row.names = F, col.names = T)
write.table(ps1 %>% tax_glom(taxrank = "Class") %>% psmelt() %>% 
        select(Class, Sample_ID, Abundance) %>% spread(Sample_ID, Abundance),
        file = "raw_abundance_tables/expr.abundance.abcls.tsv", sep = "\t", 
        quote = F, row.names = F, col.names = T)
write.table(ps1 %>% tax_glom(taxrank = "Family") %>% psmelt() %>%
        select(Family, Sample_ID, Abundance) %>% spread(Sample_ID, Abundance),
        file = "raw_abundance_tables/expr.abundance.abfam.tsv", sep = "\t", 
        quote = F, row.names = F, col.names = T)
write.table(ps1 %>% tax_glom(taxrank = "Genus") %>% psmelt() %>%
        select(Genus, Sample_ID, Abundance) %>% spread(Sample_ID, Abundance),
        file = "raw_abundance_tables/expr.abundance.abgen.tsv", sep = "\t", 
        quote = F, row.names = F, col.names = T)
```

  Create relative abundance tables (Remove Sample_ID if it doesn't exist, 
    can also change to other variables from metadata)
```{r}
ifelse(!dir.exists("relative_abundance_tables_filtered"), 
       dir.create("relative_abundance_tables_filtered"), FALSE)

write.table(ps1 %>% transform_sample_counts(function(x) {x/sum(x)} ) %>% 
              psmelt() %>% 
        arrange(OTU) %>% rename(ASV = OTU) %>% 
        select(ASV, Kingdom, Phylum, Class, Order, Family, Genus, Species, 
               Sample_ID, Abundance) %>%
        spread(Sample_ID, Abundance), 
    file = "relative_abundance_tables/expr.relative_abundance.all.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
write.table(ps1 %>% tax_glom(taxrank = "Phylum") %>% 
        transform_sample_counts(function(x) {x/sum(x)} ) %>% psmelt() %>%
        select(Phylum, Sample_ID, Abundance) %>% spread(Sample_ID, Abundance), 
    file = "relative_abundance_tables/expr.relative_abundance.abphy.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
write.table(ps1 %>% tax_glom(taxrank = "Class") %>%
        transform_sample_counts(function(x) {x/sum(x)} ) %>% psmelt() %>% 
        select(Class, Sample_ID, Abundance) %>% spread(Sample_ID, Abundance), 
    file = "relative_abundance_tables/expr.relative_abundance.abcls.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
write.table(ps1 %>% tax_glom(taxrank = "Family") %>%
        transform_sample_counts(function(x) {x/sum(x)} ) %>% psmelt() %>% 
        select(Family, Sample_ID, Abundance) %>% spread(Sample_ID, Abundance), 
    file = "relative_abundance_tables/expr.relative_abundance.abfam.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
write.table(ps1 %>% tax_glom(taxrank = "Genus") %>% 
        transform_sample_counts(function(x) {x/sum(x)} ) %>% psmelt() %>% 
        select(Genus, Sample_ID, Abundance) %>% spread(Sample_ID, Abundance), 
    file = "relative_abundance_tables/expr.relative_abundance.abgen.tsv", 
    sep = "\t", quote = F, row.names = F, col.names = T)
```
